---
title: "FDI and Growth: Panel Evidence with Income-Group Heterogeneity"
author: "Lorenzo Dufour, 800536386\\thanks{Replication materials: \\href{https://github.com/dufourlorenzo/replication-fdi-growth-panel}{github.com/dufourlorenzo/replication-fdi-growth-panel}}"
date: today
format:
  pdf:
    documentclass: article
    papersize: a4
    fontsize: 11pt
    geometry: top=1in, bottom=1in, left=1.2in, right=1.2in
    keep-tex: false
    number-sections: true
    toc: false
    colorlinks: true
    header-includes: |
      \usepackage{booktabs}
      \usepackage{float}
      \usepackage{caption}
      \usepackage{setspace}
      \usepackage{hyperref}
      \setstretch{1.2}
bibliography: references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.pos = "H", out.extra = "")
library(tidyverse)
library(fixest)
library(kableExtra)
library(WDI)
```

```{r load-data}
#| include: false
#| cache: true
# Pull all three series at once from the World Bank API.
# extra = TRUE provides the region column used to drop aggregate entries
panel_raw <- WDI(
  indicator = c(fdi    = "BX.KLT.DINV.WD.GD.ZS",
                gdppcg = "NY.GDP.PCAP.KD.ZG",
                gdppc  = "NY.GDP.PCAP.PP.KD"),
  country   = "all",
  extra     = TRUE
) |>
  filter(region != "Aggregates", !is.na(iso3c), iso3c != "") |>
  # for reproducibility, restrict to the period covered to the one available at the time of analysis
  filter(year >= 1960, year <= 2025) |> 
  rename(code = iso3c) |>
  select(country, code, year, fdi, gdppcg, gdppc)

# Characterise the raw FDI series before any filtering.
# Computed here because zeros/negatives often coincide with missing GDP data
# and would be lost in the complete-cases filter below.
pct_zero <- round(100 * mean(panel_raw$fdi == 0, na.rm = TRUE), 1)
pct_neg  <- round(100 * mean(panel_raw$fdi < 0,  na.rm = TRUE), 1)

panel_raw <- panel_raw |>
  filter(!is.na(fdi), !is.na(gdppcg), !is.na(gdppc), gdppc > 0) |>
  mutate(lgdppc = log(gdppc))

# Winsorise FDI before lagging so the percentiles come from the raw distribution
q_f       <- quantile(panel_raw$fdi, c(0.01, 0.99), na.rm = TRUE)
panel_raw <- mutate(panel_raw, fdi = pmin(pmax(fdi, q_f[1]), q_f[2]))

panel <- panel_raw |>
  group_by(code) |>
  arrange(year) |>
  mutate(fdi_l1    = lag(fdi, 1),
         fdi_l2    = lag(fdi, 2),
         fdi_l3    = lag(fdi, 3),
         lgdppc_l1 = lag(lgdppc, 1)) |>
  ungroup() |>
  filter(!is.na(fdi_l1), !is.na(lgdppc_l1))

# Winsorise growth after lagging so percentiles cover the estimation sample
q_g   <- quantile(panel$gdppcg, c(0.01, 0.99), na.rm = TRUE)
panel <- mutate(panel, gdppcg = pmin(pmax(gdppcg, q_g[1]), q_g[2]))

fdi_w_lo <- round(q_f[1], 1)
fdi_w_hi <- round(q_f[2], 1)
```

```{r analysis}
#| include: false
# Classify countries by time-averaged PPP GDP per capita vs. sample median
med_gdppc   <- panel |>
  group_by(code) |>
  summarise(avg = mean(gdppc, na.rm = TRUE), .groups = "drop") |>
  pull(avg) |>
  median()
med_gdppc_k <- round(med_gdppc / 1000)

est <- panel |>
  mutate(low_inc    = as.integer(mean(gdppc, na.rm = TRUE) < med_gdppc),
         .by = code) |>
  mutate(ihs_fdi    = asinh(fdi_l1),
         ihs_fdi_li = ihs_fdi * low_inc)

# Time-varying income classification (robustness R1): below-median in each year
yr_med <- est |>
  group_by(year) |>
  summarise(yr_med_gdppc = median(gdppc, na.rm = TRUE), .groups = "drop")

est <- est |>
  left_join(yr_med, by = "year") |>
  mutate(low_inc_tv    = as.integer(gdppc < yr_med_gdppc),
         ihs_fdi_li_tv = ihs_fdi * low_inc_tv)

# Balanced sub-panel (robustness R2)
min_obs_bal     <- 20
est_bal         <- est |> group_by(code) |> filter(n() >= min_obs_bal) |> ungroup()
n_countries_bal <- n_distinct(est_bal$code)

n_countries  <- n_distinct(est$code)
n_obs        <- nrow(est)
yr_min       <- min(est$year)
yr_max       <- max(est$year)
n_lowinc     <- n_distinct(est$code[est$low_inc == 1])
mean_grow    <- round(mean(est$gdppcg, na.rm = TRUE), 1)
sd_grow      <- round(sd(est$gdppcg, na.rm = TRUE), 1)
mean_ihs     <- round(mean(est$ihs_fdi, na.rm = TRUE), 1)
max_ihs      <- round(max(est$ihs_fdi, na.rm = TRUE), 1)
mean_lgdppc  <- round(mean(est$lgdppc_l1, na.rm = TRUE), 2)
mean_gdppc_k <- round(exp(mean(est$lgdppc_l1, na.rm = TRUE)) / 1000, 1)
mean_grow_li <- round(mean(est$gdppcg[est$low_inc == 1], na.rm = TRUE), 1)

setFixest_dict(c(ihs_fdi       = "asinh FDI(t-1)",
                 ihs_fdi_li    = "asinh FDI(t-1) \u00d7 Below-median",
                 ihs_fdi_li_tv = "asinh FDI(t-1) \u00d7 Below-median (TV)",
                 lgdppc_l1     = "ln GDPpc(t-1)",
                 gdppcg        = "GDPpc Growth"))

m1 <- feols(gdppcg ~ ihs_fdi,                                          data = est, cluster = ~code)
m2 <- feols(gdppcg ~ ihs_fdi | code + year,                            data = est, cluster = ~code)
m3 <- feols(gdppcg ~ ihs_fdi + lgdppc_l1 | code + year,               data = est, cluster = ~code)
m4 <- feols(gdppcg ~ ihs_fdi + ihs_fdi_li + lgdppc_l1 | code + year,  data = est, cluster = ~code)

b1        <- coef(m1)["ihs_fdi"]
b2        <- coef(m2)["ihs_fdi"]
b3        <- coef(m3)["ihs_fdi"]
d3        <- coef(m3)["lgdppc_l1"]
b4_h      <- coef(m4)["ihs_fdi"]
b4_l      <- coef(m4)["ihs_fdi"] + coef(m4)["ihs_fdi_li"]
d4        <- coef(m4)["lgdppc_l1"]
pct_reduc  <- round(100 * (b1 - b2) / b1)
med_fdi_li <- median(est$fdi_l1[est$low_inc == 1], na.rm = TRUE)
marg       <- b4_l / sqrt(1 + med_fdi_li^2)

get_p <- function(m, var) coeftable(m)[var, "Pr(>|t|)"]

m_tv   <- feols(gdppcg ~ ihs_fdi + ihs_fdi_li_tv + lgdppc_l1 | code + year, data = est,     cluster = ~code)
m_bal  <- feols(gdppcg ~ ihs_fdi + ihs_fdi_li    + lgdppc_l1 | code + year, data = est_bal, cluster = ~code)

b_tv_h  <- coef(m_tv)["ihs_fdi"]
b_tv_l  <- coef(m_tv)["ihs_fdi"] + coef(m_tv)["ihs_fdi_li_tv"]
p_tv_h  <- get_p(m_tv, "ihs_fdi")
b_bal_h <- coef(m_bal)["ihs_fdi"]
b_bal_l <- coef(m_bal)["ihs_fdi"] + coef(m_bal)["ihs_fdi_li"]
p_bal_h <- get_p(m_bal, "ihs_fdi")

est_l2 <- est |> filter(!is.na(fdi_l2)) |>
  mutate(ihs_fdi_l2    = asinh(fdi_l2),
         ihs_fdi_l2_li = asinh(fdi_l2) * low_inc)
est_l3 <- est |> filter(!is.na(fdi_l3)) |>
  mutate(ihs_fdi_l3    = asinh(fdi_l3),
         ihs_fdi_l3_li = asinh(fdi_l3) * low_inc)

m_l2 <- feols(gdppcg ~ ihs_fdi_l2 + ihs_fdi_l2_li + lgdppc_l1 | code + year, data = est_l2, cluster = ~code)
m_l3 <- feols(gdppcg ~ ihs_fdi_l3 + ihs_fdi_l3_li + lgdppc_l1 | code + year, data = est_l3, cluster = ~code)

b_l2_h   <- coef(m_l2)["ihs_fdi_l2"]
b_l2_l   <- coef(m_l2)["ihs_fdi_l2"] + coef(m_l2)["ihs_fdi_l2_li"]
p_l2_h   <- get_p(m_l2, "ihs_fdi_l2")
n_obs_l2 <- nobs(m_l2)
b_l3_h   <- coef(m_l3)["ihs_fdi_l3"]
b_l3_l   <- coef(m_l3)["ihs_fdi_l3"] + coef(m_l3)["ihs_fdi_l3_li"]
p_l3_h   <- get_p(m_l3, "ihs_fdi_l3")
n_obs_l3 <- nobs(m_l3)

p4_h_2w  <- pvalue(m4, vcov = ~code + year)["ihs_fdi"]
p4_i_2w  <- pvalue(m4, vcov = ~code + year)["ihs_fdi_li"]
se4_h_2w <- se(m4, vcov = ~code + year)["ihs_fdi"]
se4_i_2w <- se(m4, vcov = ~code + year)["ihs_fdi_li"]

# Wald test for the total below-median FDI effect (b1 + b2) under any vcov
total_li_p <- function(m, fdi_var = "ihs_fdi", int_var = "ihs_fdi_li", vcov_spec = ~code) {
  V  <- vcov(m, vcov = vcov_spec)
  b  <- coef(m)[fdi_var] + coef(m)[int_var]
  se <- sqrt(V[fdi_var, fdi_var] + V[int_var, int_var] + 2 * V[fdi_var, int_var])
  2 * pnorm(-abs(b / se))
}

p4_l_1w <- total_li_p(m4)
p4_l_2w <- total_li_p(m4, vcov_spec = ~code + year)
p_tv_l  <- total_li_p(m_tv,  int_var = "ihs_fdi_li_tv")
p_bal_l <- total_li_p(m_bal)
p_l2_l  <- total_li_p(m_l2, fdi_var = "ihs_fdi_l2", int_var = "ihs_fdi_l2_li")
p_l3_l  <- total_li_p(m_l3, fdi_var = "ihs_fdi_l3", int_var = "ihs_fdi_l3_li")

stars_p <- function(p) {
  if      (p < 0.01) "***"
  else if (p < 0.05) "**"
  else if (p < 0.10) "*"
  else               ""
}

sig_label <- function(p) {
  if      (p < 0.001) "$p < 0.001$"
  else if (p < 0.01)  "$p < 0.01$"
  else if (p < 0.05)  "$p < 0.05$"
  else if (p < 0.10)  "$p < 0.10$"
  else                "n.s."
}
```

```{r abstract, results='asis'}
cat(paste0(
  "\\begin{abstract}\n",
  "This paper examines the relationship between foreign direct investment (FDI) and ",
  "host-country economic growth using an unbalanced panel of ", n_countries,
  " economies over the period ", yr_min, " to ", yr_max, ". ",
  "To mitigate reverse causality, FDI is lagged by one period and two-way fixed effects are employed ",
  "that absorb time-invariant country characteristics and global time trends. ",
  "A conditional convergence control following Barro (1991) further reduces omitted-variable bias. ",
  "The main contribution is an income-group interaction that tests Findlay's (1978) ",
  "relative-backwardness hypothesis: whether technology spillovers from FDI are larger ",
  "where the gap with the global frontier is greatest. ",
  "The results suggest that the aggregate FDI-growth relationship is attenuated once fixed ",
  "effects are introduced, but masks income-group heterogeneity. ",
  "In above-median-income economies the FDI coefficient is statistically indistinguishable from zero. ",
  "In below-median-income economies a one-percentage-point increase in the lagged FDI-to-GDP ratio is associated with approximately ",
  round(marg, 2),
  " percentage points of additional annual GDP per capita growth, ",
  "significant at the ",
  if(p4_l_1w < 0.001) "0.1\\%" else if(p4_l_1w < 0.01) "1\\%" else if(p4_l_1w < 0.05) "5\\%" else if(p4_l_1w < 0.10) "10\\%" else "conventional",
  " level. ",
  "These findings are consistent with technology-transfer mechanisms and have ",
  "implications for development policy.\n",
  "\\end{abstract}\n"
))
```

# Introduction

In recent decades, governments worldwide have competed to attract foreign direct investment (FDI), offering tax incentives, special economic zones, and investment protection treaties in the belief that inward capital generates jobs, transfers technology, and accelerates economic growth. Yet the empirical evidence on whether FDI delivers these gains remains contested. While the literature documents positive correlations between FDI and various development outcomes, establishing causality is harder: faster-growing, more stable economies also attract more FDI, potentially reversing the direction of causation.

The distinction between these two interpretations has policy consequences. The case of Costa Rica is cited as an illustration of FDI-led structural transformation: from the 1990s onward, the country attracted FDI inflows in manufacturing and services, and this experience is often held up as evidence for the potential of export-oriented FDI [@voxdev_costarica]. Whether FDI drove that transformation or reflected Costa Rica's prior investments in human capital and institutional stability is disputed. If FDI causes growth, investment promotion agencies represent an efficient use of public resources; if FDI is primarily reactive, those resources may be better directed at improving fundamentals.

This paper contributes evidence by exploiting within-country temporal variation in an unbalanced panel covering `r n_countries` economies from `r yr_min` to `r yr_max`. Three methodological choices distinguish the approach. First, FDI is lagged by one period relative to growth, which mitigates the most direct form of simultaneity bias. Second, two-way fixed effects are employed, absorbing time-invariant country characteristics (geography, colonial history, institutional quality) and year-specific global shocks, so that identification draws on within-country changes in FDI over time. Third, the analysis tests for income-group heterogeneity in the FDI-growth relationship. The theoretical prior, dating to @findlay1978, is that technology spillovers from FDI are largest where the technology gap between the host economy and the global frontier is greatest, precisely in lower-income countries. Testing this mechanism requires both a long panel and the income-group interaction employed here.

The main results are as follows. The pooled cross-sectional correlation between FDI and growth is positive and large, confirming earlier findings in the literature. Once country and year fixed effects are introduced, however, the average effect falls, consistent with level selection (wealthier countries attracting more FDI for reasons unrelated to FDI's own effect on growth). For above-median-income economies, the fixed-effects estimate is statistically indistinguishable from zero. For below-median-income economies the total effect is approximately `r round(b4_l, 2)` percentage points of GDP per capita growth per unit of asinh(FDI) (`r sig_label(p4_l_1w)`). These findings are consistent with @findlay1978's relative-backwardness hypothesis and with firm-level evidence from @pasali2020.

The remainder of the paper is organised as follows. Section 2 reviews the relevant theoretical and empirical literature. Section 3 describes the data. Section 4 lays out the empirical strategy. Section 5 presents and discusses the results. Section 6 concludes with policy implications and directions for future research.

# Background

## Theoretical Foundations

The economics of FDI and growth has been shaped by three main theoretical traditions. The first concerns the motivations for multinational activity. @dunning1979 synthesises earlier contributions in the "OLI" (Ownership-Location-Internalisation) eclectic framework: a firm invests abroad when it holds firm-specific ownership advantages (technology, brand, management know-how), when exploiting them in a foreign location offers productivity or market benefits, and when internal organisation of cross-border transactions is cheaper than licensing. The OLI paradigm emphasises that FDI is primarily a vehicle for firms to exploit, not develop, their advantages; yet the host country may benefit from knowledge spillovers as local firms and workers interact with foreign affiliates.

Two theoretical frameworks motivate the income-heterogeneity prediction tested in this paper. @findlay1978 derives that the rate of technological catching-up in the host country is increasing in the technology gap between the foreign and domestic sectors: countries furthest from the global frontier should experience the largest productivity gains from each unit of inward FDI, a result known as the "relative-backwardness hypothesis." @wangblomstrom1992 model technology diffusion as a function of the multinational's investment in transferring its technology stock and the host economy's absorptive capacity. Their framework implies that spillovers are attenuated when absorptive capacity is too low, suggesting gains accrue primarily to economies above a minimum absorptive-capacity threshold. These two mechanisms generate a tension: Findlay predicts larger gains for lower-income (higher technology-gap) countries, while Wang and Blomström caution that the lowest-income economies may lack the absorptive capacity to capture those gains. The income-group interaction estimated below provides a reduced-form test of whether below-median-income economies in the aggregate benefit more from FDI than above-median-income economies; it cannot cleanly distinguish between the two mechanisms and should be read as a test of the net effect of income on FDI-growth sensitivity.

The growth literature provides the second building block. @barro1991's cross-country evidence on conditional convergence establishes that, conditional on structural characteristics, poorer countries grow faster, consistent with neoclassical diminishing returns to capital. This motivates including initial income as a control variable: without it, the negative correlation between income levels and growth would confound any variable that is itself correlated with income. In the two-way fixed-effects specification, the log of lagged GDP per capita proxies for a country's distance from its own steady state, and the recovered coefficient ($\hat\delta \approx `r round(d3, 2)`$) provides a useful specification check. Note that in a within-country fixed-effects framework this coefficient captures within-country mean reversion rather than the cross-country convergence documented by @barro1991; the negative sign is expected under the Solow mechanism but should not be overinterpreted as confirming it.

## Empirical Literature

The empirical literature on FDI and growth is large and yields mixed results. @kechagia2018 survey sixty years of empirical research spanning 1950 to 2015, documenting a wide range of findings: while many studies identify positive effects, a share find negligible or negative relationships, and magnitudes vary considerably with sample, period, and identification strategy. Several mechanisms have been proposed to explain this heterogeneity: absorptive capacity (countries below a human capital or financial development threshold cannot translate FDI into growth), the nature of FDI (resource-seeking vs. efficiency-seeking), and the direction of causation. @borensztein1998 find that FDI raises growth only in countries with sufficient human capital, providing early cross-country evidence for an absorptive-capacity threshold. @alfaro2004 show that local financial market development is a conditioning factor: FDI has a negligible or negative effect on growth in countries with underdeveloped financial systems, but a positive effect where domestic credit markets are deep. @carkovic2005 apply GMM estimators to a large cross-country panel and find little evidence of an independent causal effect of FDI on growth once country fixed effects and other growth determinants are controlled for, a finding that underscores the identification challenge.

@narula2012 raise a further challenge to the causal interpretation of positive FDI-growth correlations. In their introductory essay to a debate forum on FDI and development, they argue that the empirical evidence on FDI-assisted growth is more ambiguous than is widely acknowledged, and that structural simplifications in the empirical literature obscure the conditions under which FDI contributes to host-country growth. They point to reverse causality (investors targeting already-dynamic economies), heterogeneous spillover conditions, and the inadequacy of aggregate FDI data as persistent sources of bias. The approach taken here addresses the reverse-causality concern through lagged regressors and fixed effects, but, as discussed in Section 5.4, a valid instrumental-variable strategy would be needed to fully resolve the endogeneity problem.

@blomstrom1991 surveys the evidence on FDI spillovers, cataloguing the channels (competitive pressure, labour training, and technology transfer to local suppliers) through which host-country firms may benefit from multinational presence. The survey notes that spillover gains are likely larger where local firms have the skill base to absorb and imitate foreign technologies, anticipating the absorptive-capacity literature. Firm-level evidence from @pasali2020 suggests that productivity gains from foreign ownership are concentrated among smaller firms in developing countries, consistent with the macro-level heterogeneity estimated in this paper. @tomohara2011 document positive wage effects of FDI in manufacturing, consistent with labour-market spillovers in a developing-country context.

## Contribution

Against this backdrop, the present paper makes two contributions. Methodologically, it applies a two-way fixed-effects estimator with country-clustered standard errors to a panel of `r n_countries` economies, which is more demanding than the pooled or cross-section designs common in the earlier literature. Substantively, it tests the @findlay1978 heterogeneity prediction using an income-group interaction in a within-country identification framework, rather than relying on cross-sectional variation in income that is confounded with many other country characteristics. The main finding is that the FDI-growth relationship is statistically indistinguishable from zero for above-median-income economies and positive for below-median-income economies, consistent with the relative-backwardness hypothesis, though the identification strategy falls short of the IV benchmark required for a causal interpretation.

# Data

## Sources and Variables

The analysis draws on three annual series from the World Bank Development Indicators [@worldbank2026]:

1. FDI net inflows as a percentage of GDP (series `BX.KLT.DINV.WD.GD.ZS`): the explanatory variable.
2. GDP per capita growth, annual % (series `NY.GDP.PCAP.KD.ZG`): the dependent variable.
3. GDP per capita, PPP (constant 2021 international $) (series `NY.GDP.PCAP.PP.KD`): used to construct the convergence control and the income-group indicator.

The PPP-adjusted constant-price series is preferred over nominal current-USD GDP per capita for two reasons. First, PPP adjustment removes cross-country price-level differences, ensuring that the income-group classification reflects real purchasing power rather than nominal exchange rates. Second, using constant prices removes inflation effects, so that the convergence control and income-group indicator reflect real income changes over time.

After removing regional and income-group aggregates from the World Bank's reporting structure, the raw data cover `r n_countries` individual economies. The panel is unbalanced because FDI reporting begins at different dates for different countries; the earliest observations are from `r yr_min` and the latest from `r yr_max`.

## Data Cleaning

All estimation uses R [@rcoreteam2025] with packages WDI [@arelbundock2019wdi], fixest [@berge2018fixest], tidyverse [@wickham2019tidyverse], and kableExtra [@zhu2024kableextra].

Three decisions require explicit justification. First, both the dependent variable and the FDI regressor are winsorised at the 1st and 99th percentiles of their respective pooled distributions. The raw growth series contains episodes of hyperinflation, armed conflict, and statistical revisions that generate observations ranging below −50% and above 100%. FDI-to-GDP ratios can exceed 1,000% in small economies that serve as holding-company hubs (e.g. the Cayman Islands), reflecting balance-sheet intermediation rather than real productive investment. Winsorising growth to [`r round(q_g[1],1)`%, `r round(q_g[2],1)`%] and FDI to [`r fdi_w_lo`%, `r fdi_w_hi`%] bounds these extremes while retaining all observations; no country is excluded by list or name. This treatment is applied to the raw annual FDI series before constructing lags, so the IHS transformation is then applied to the already-winsorised lagged values.

Second, the FDI series contains `r pct_zero`% zeros and `r pct_neg`% negative values. Zeros arise in years when no investment inflows were recorded (common in countries with thin financial sectors), while negative values reflect disinvestment and profit repatriation. The natural logarithm cannot handle these values: $\log(x)$ is undefined for $x \leq 0$. Dropping zeros and negatives would remove precisely the low-FDI, often lower-income observations most relevant to the heterogeneity analysis, inducing selection bias. The inverse hyperbolic sine (IHS) transformation $\text{asinh}(x) = \ln(x + \sqrt{x^2+1})$ avoids this problem: it is defined for all real $x$, equals zero when $x = 0$, handles negative values symmetrically, and approximates $\ln(2x)$ for large positive $x$ [@bellemare2020]. Coefficients on $\text{asinh}(\text{FDI})$ are interpretable as semi-elasticities, similarly to log-transformed regressors.

Third, the income-group indicator $D_i^L$ classifies each country as below- or above-median by comparing its time-averaged GDP per capita to the cross-country median of those averages. This produces `r n_lowinc` below-median economies out of `r n_countries` total, with a median threshold of approximately \$`r med_gdppc_k`k in constant 2021 international dollars (PPP). The time-averaging removes cyclical variation and classifies countries by their structural income level rather than their position in any given year.

## Summary Statistics

```{r tbl-desc}
#| tbl-cap: "Summary Statistics: Estimation Sample"

vars <- est |>
  select(gdppcg, ihs_fdi, lgdppc_l1) |>
  rename(`GDPpc Growth (%)` = gdppcg,
         `asinh(FDI), lagged` = ihs_fdi,
         `Log GDPpc, lagged`  = lgdppc_l1)

tbl <- vars |>
  summarise(across(everything(),
                   list(Mean = ~round(mean(., na.rm=TRUE), 2),
                        SD   = ~round(sd(.,   na.rm=TRUE), 2),
                        Min  = ~round(min(.,  na.rm=TRUE), 2),
                        Max  = ~round(max(.,  na.rm=TRUE), 2),
                        N    = ~sum(!is.na(.))))) |>
  pivot_longer(everything(), names_to = c("Variable","Stat"),
               names_sep = "_(?=[^_]+$)") |>
  pivot_wider(names_from = Stat, values_from = value)

kbl(tbl, format = "latex", booktabs = TRUE,
    col.names = c("Variable","Mean","SD","Min","Max","N"),
    caption   = "Summary statistics for the three core variables in the estimation sample. Both growth and FDI are winsorised at the 1st and 99th percentiles before model estimation. FDI is reported after the IHS transformation is applied to the winsorised lagged raw value.") |>
  kable_styling(latex_options = c("hold_position","scale_down"), font_size = 9)
```

Table 1 shows that the average GDP per capita growth rate in the sample is `r mean_grow`%, with a standard deviation of `r sd_grow` percentage points after winsorisation. The IHS-transformed FDI variable has a mean of `r mean_ihs` and a maximum of `r max_ihs`. The mean log GDP per capita of `r mean_lgdppc` corresponds to approximately \$`r mean_gdppc_k`k in constant 2021 international dollars (PPP).

```{r fig-scatter, fig.width=5.5, fig.height=2.8}
#| fig-cap: "IHS-transformed lagged FDI (x-axis) and GDP per capita growth (y-axis). Both variables are winsorised at the 1st and 99th percentiles. Lines are separate OLS fits by income group."

est |>
  mutate(group = if_else(low_inc == 1, "Below-median income", "Above-median income")) |>
  ggplot(aes(x = ihs_fdi, y = gdppcg, colour = group)) +
  geom_point(alpha = 0.08, size = 0.5) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.9) +
  scale_colour_manual(values = c("Above-median income" = "steelblue",
                                 "Below-median income" = "firebrick")) +
  labs(x = expression(asinh(FDI[t-1]~"(% GDP)")),
       y = "GDPpc Growth (%)",
       colour = NULL) +
  theme_classic(base_size = 9) +
  theme(legend.position = "bottom")
```

Figure 1 plots the raw relationship. A steeper gradient is visible for below-median-income countries, consistent with the heterogeneity that motivates the main specification.

# Methodology

## Identification Challenges

Two main threats complicate causal inference. The first is reverse causality: multinational investors target growing, politically stable economies, so that a positive FDI-growth correlation may reflect growth causing FDI rather than FDI causing growth. One-period lagging of the regressor mitigates this concern without eliminating it entirely.

The second challenge is omitted variables. Countries that attract more FDI may also implement institutional reforms, liberalise trade, or invest in infrastructure in ways that independently raise growth. Unobservable time-invariant characteristics (geography, legal origins, colonial history) may simultaneously attract FDI and correlate with long-run growth levels. Country fixed effects absorb all such time-invariant confounders. Year fixed effects remove any global shock (commodity price cycles, global financial crises, pandemic effects) that affects all countries simultaneously and is correlated with both global FDI flows and global growth.

A third, less easily addressed threat is time-varying omitted variables. Country fixed effects absorb any determinant of growth that is constant within a country, but not variables that change over time: trade liberalisation episodes, domestic financial deepening, institutional reform, or infrastructure investment may all be correlated with FDI inflows and independently affect growth. The specification includes only one time-varying control (lagged income) beyond FDI, so time-varying confounders remain a residual concern. This is an acknowledged limitation of the design, and results should be interpreted as conditional associations rather than causal estimates.

## Model Specifications

Specification 1: Pooled OLS Baseline. The most transparent starting point is a pooled regression with no controls beyond the lagged FDI variable:

$$g_{it} = \alpha + \beta_1\,\text{asinh}(\text{FDI}_{i,t-1}) + \varepsilon_{it} \tag{1}$$

where $g_{it}$ denotes annual GDP per capita growth for country $i$ in year $t$. This benchmark captures the unconditional correlation and is comparable to the bulk of the early literature.

Specification 2: Two-Way Fixed Effects. To exploit within-country temporal variation and control for omitted country-level factors, country ($\alpha_i$) and year ($\lambda_t$) fixed effects are added:

$$g_{it} = \alpha_i + \lambda_t + \beta_1\,\text{asinh}(\text{FDI}_{i,t-1}) + \varepsilon_{it} \tag{2}$$

The coefficient $\hat\beta_1$ is identified from within-country deviations: the estimator removes each country's time mean and each year's cross-country mean, so variation in FDI that is common to all years for a given country (level differences) and variation common to all countries in a given year (global cycles) are both absorbed. What remains is the idiosyncratic change in a country's FDI position relative to its own history, net of the common year effect. This sweeps out confounding from any time-invariant determinant of growth.

Specification 3: Conditional Convergence Control. Following @barro1991, the log of lagged GDP per capita is added as a control for conditional convergence:

$$g_{it} = \alpha_i + \lambda_t + \beta_1\,\text{asinh}(\text{FDI}_{i,t-1}) + \delta\ln\text{GDPpc}_{i,t-1} + \varepsilon_{it} \tag{3}$$

A negative $\hat\delta$ would confirm that, conditional on fixed effects, countries further from their steady state grow faster, a standard check that the specification captures the convergence mechanism documented in the growth literature.

Specification 4: Income-Group Heterogeneity. The primary specification augments Equation 3 with an interaction term:

$$g_{it} = \alpha_i + \lambda_t + \beta_1\,\text{asinh}(\text{FDI}_{i,t-1}) + \beta_2\bigl[\text{asinh}(\text{FDI}_{i,t-1})\times D_i^L\bigr] + \delta\ln\text{GDPpc}_{i,t-1} + \varepsilon_{it} \tag{4}$$

where $D_i^L = 1$ if country $i$'s time-averaged GDP per capita falls below the cross-country sample median. The country fixed effects absorb the (time-invariant) main effect of $D_i^L$. The interaction coefficient $\hat\beta_2$ captures the additional FDI-growth sensitivity of below-median-income economies relative to above-median-income economies within the same panel. Under the relative-backwardness hypothesis [@findlay1978], $\hat\beta_2 > 0$, though it should be noted that the within-country identification cannot fully isolate the Findlay mechanism from other income-related heterogeneity.

Standard errors are clustered by country throughout, accounting for arbitrary serial correlation within the country time series. Section 5.4 reports four robustness checks that directly address the limitations of this baseline: (i) a time-varying income classification that re-classifies countries each year against the annual cross-country median (addressing the convergence-country misclassification concern); (ii) a balanced panel restricting to countries with at least `r min_obs_bal` annual observations; (iii) alternative lag structures at $t-2$ and $t-3$; and (iv) two-way country-and-year clustered standard errors.

# Results

## Regression Estimates

```{r tbl-regs, results='asis'}
#| tbl-cap: "FDI and GDP Per Capita Growth: Panel Estimates"

etable(m1, m2, m3, m4,
       tex        = TRUE,
       se         = "cluster",
       cluster    = ~code,
       signif.code = c("***" = 0.01, "**" = 0.05, "*" = 0.10),
       headers    = list("Pooled OLS" = 1, "Two-way FE" = 3),
       drop       = "Intercept",
       fitstat    = c("n","r2"),
       digits     = 3,
       notes      = "Country-clustered standard errors in parentheses. * p<0.10, ** p<0.05, *** p<0.01. Below-median dummy equals 1 if country's time-averaged GDP per capita (PPP, constant 2021 international $) is below the cross-country sample median. Country and year fixed effects absorbed in columns (2) to (4).")
```

Table 2 presents the four specifications, discussed in turn below.

Column (1): Pooled OLS. The unconditional association is positive: $\hat\beta_1 = `r round(b1, 3)`$ (`r sig_label(get_p(m1, "ihs_fdi"))`), implying that a one-unit increase in $\text{asinh}(\text{FDI})$ is associated with `r round(b1, 2)` percentage points of additional annual growth. This is consistent with the positive correlations reported in the survey literature [@kechagia2018]. This estimate likely overstates any causal effect due to omitted country heterogeneity.

Column (2): Two-Way FE without convergence control. Adding country and year fixed effects reduces the estimate to $\hat\beta_1 = `r round(b2, 3)`$, still statistically significant. The `r pct_reduc`% reduction from column (1) is consistent with a level-selection effect: countries with higher FDI tend to be wealthier and to grow faster for reasons unrelated to FDI.

Column (3): Two-Way FE with convergence control. The addition of $\ln\text{GDPpc}_{i,t-1}$ leaves the FDI coefficient unchanged at `r round(b3, 3)`. The convergence coefficient $\hat\delta \approx `r round(d3, 2)`$ (`r sig_label(get_p(m3, "lgdppc_l1"))`) is negative: a 10% higher initial income is associated with approximately `r round(abs(d3) * log(1.1), 2)` pp less annual growth, conditional on fixed effects, consistent with @barro1991.

Column (4): Income-group heterogeneity. The results differ across income groups. For above-median-income economies, the FDI semi-elasticity falls to $\hat\beta_1 = `r round(b4_h, 3)`$, which is statistically indistinguishable from zero. For below-median-income economies the total effect is $\hat\beta_1 + \hat\beta_2 \approx `r round(b4_l, 3)`$ (`r sig_label(p4_l_1w)`). The positive $\hat\beta_2$ is consistent with the pattern predicted by @findlay1978.

## Economic Magnitude

To translate the econometric estimate into a more interpretable magnitude, the marginal effect is evaluated at the median lagged FDI-to-GDP ratio among below-median-income countries in the sample. Since the regressor is $\text{asinh}(\text{FDI}_{i,t-1})$, the derivative with respect to the lagged FDI share is $d\,\text{asinh}(x)/dx = 1/\sqrt{1+x^2}$, so the implied effect in the original percentage-of-GDP space is:

$$\frac{\partial g_{it}}{\partial \text{FDI}_{i,t-1}} \approx \frac{\hat\beta_1 + \hat\beta_2}{\sqrt{1 + \text{FDI}^2_{\text{med}}}} \approx `r round(marg, 3)`$$

evaluated at the median FDI share of `r round(med_fdi_li, 1)`% for below-median-income countries. A one-percentage-point increase in the lagged FDI-to-GDP ratio is therefore associated with approximately `r round(marg, 2)` pp more annual GDP per capita growth in below-median-income economies. The mean growth rate in this group is `r mean_grow_li`%, so the implied effect represents about `r round(100 * marg / mean_grow_li)`% of mean growth.

## Consistency with Theory and Micro Evidence

The income-group heterogeneity in column (4) is consistent with the @findlay1978 argument at the macroeconomic level and with the micro-level evidence in @pasali2020, who find that firm-level productivity gains from foreign ownership are concentrated among smaller firms in developing countries, consistent with firms facing a larger technology gap relative to multinational affiliates; this provides micro-foundations for the macro-level heterogeneity estimated here. A second complementary mechanism is capital scarcity: in below-median-income economies with shallow domestic financial systems, FDI can augment the productive capital stock directly, beyond any knowledge-diffusion channel [@dunning1979]. Both mechanisms are less likely in above-median-income countries, which operate closer to the technology frontier and have deeper capital markets. One possible explanation for the statistically indistinguishable-from-zero effect in above-median-income economies is competitive displacement (multinational affiliates crowding out domestic firms), though the design cannot distinguish this from simply lower technology-gap gains; this remains a question for future research using sector-level data.

The result that adding the income interaction leaves the convergence coefficient at $\hat\delta \approx `r round(d4, 2)`$ (`r sig_label(get_p(m4, "lgdppc_l1"))`), nearly identical to column (3), is consistent with the interaction and the convergence control capturing distinct sources of variation rather than the same mechanism.

## Robustness Checks

```{r tbl-robust, results='asis'}
#| tbl-cap: "Robustness Checks: Income-Group FDI Effects across Alternative Specifications"

rob <- tibble::tibble(
  Check = c(
    "Baseline (m4): static income groups, country-clustered SE",
    "R1: Time-varying income classification (annual median)",
    paste0("R2: Balanced panel ($\\geq ", min_obs_bal, "$ obs/country; $n_{\\text{cty}}=", n_countries_bal, "$)"),
    "R3a: Two-year lag, asinh(FDI$_{t-2}$)",
    "R3b: Three-year lag, asinh(FDI$_{t-3}$)",
    "R4: Two-way clustered SE (country $\\times$ year)"
  ),
  `$N$` = format(c(
    n_obs,
    nobs(m_tv),
    nobs(m_bal),
    n_obs_l2,
    n_obs_l3,
    n_obs
  ), big.mark = ","),
  `Above-median $\\hat\\beta$` = c(
    sprintf("%.3f%s", b4_h,   stars_p(get_p(m4,  "ihs_fdi"))),
    sprintf("%.3f%s", b_tv_h, stars_p(p_tv_h)),
    sprintf("%.3f%s", b_bal_h,stars_p(p_bal_h)),
    sprintf("%.3f%s", b_l2_h, stars_p(p_l2_h)),
    sprintf("%.3f%s", b_l3_h, stars_p(p_l3_h)),
    sprintf("%.3f%s", b4_h,   stars_p(p4_h_2w))
  ),
  `Below-median total` = c(
    sprintf("%.3f%s", b4_l,   stars_p(p4_l_1w)),
    sprintf("%.3f%s", b_tv_l, stars_p(p_tv_l)),
    sprintf("%.3f%s", b_bal_l,stars_p(p_bal_l)),
    sprintf("%.3f%s", b_l2_l, stars_p(p_l2_l)),
    sprintf("%.3f%s", b_l3_l, stars_p(p_l3_l)),
    sprintf("%.3f%s", b4_l,   stars_p(p4_l_2w))
  )
)

kbl(rob, format = "latex", booktabs = TRUE, escape = FALSE,
    col.names = c("Specification", "$N$",
                  "Above-median $\\hat\\beta$",
                  "Below-median total effect"),
    caption = paste0(
      "Robustness of the income-heterogeneity result. ",
      "Stars on both columns: *p$<$0.10, **p$<$0.05, ***p$<$0.01 (country-clustered SE). ",
      "Above-median $\\hat\\beta$ is the coefficient on asinh(FDI$_{t-k}$); ",
      "Below-median total effect is $\\hat\\beta_1+\\hat\\beta_2$ tested via Wald test. ",
      "R4 uses same point estimates as the baseline but reclusters SE by country and year simultaneously. ",
      "$N$ for R3a and R3b are smaller because the additional lag removes the first one or two years per country.")) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9) |>
  column_spec(1, width = "7cm") |>
  row_spec(0, bold = TRUE)
```

Table 3 summarises four robustness exercises (R1-R4), the third of which comprises two lag variants (t-2 and t-3). The baseline result is qualitatively unchanged across all four specifications.

R1: Time-varying income classification. The static $D_i^L$ dummy assigns countries to a group for their entire history, potentially misclassifying rapidly converging economies (e.g., South Korea, Singapore) in later decades. To address this, $D_i^L$ is replaced with a time-varying indicator $D_{it}^L = \mathbf{1}[\text{GDPpc}_{it} < \text{median}_j(\text{GDPpc}_{jt})]$ that re-classifies each country against the annual cross-country median. The total below-median-income effect under this specification is $`r round(b_tv_l, 3)`$ (`r sig_label(p_tv_l)`), smaller in magnitude than the baseline ($`r round(b4_l, 3)`$) but still positive and statistically significant, suggesting the result is not an artefact of time-fixed group assignment. The attenuation is consistent with the time-varying classification more aggressively reassigning converging countries to the above-median group in later periods.

R2: Balanced panel. The main sample is an unbalanced panel: countries with short data histories contribute only a handful of observations and could disproportionately drive the results. To assess this, the sample is restricted to the `r n_countries_bal` countries with at least `r min_obs_bal` annual observations, ensuring that each country's fixed effect is estimated from a meaningful number of time periods. The total below-median-income effect under this restriction is $`r round(b_bal_l, 3)`$ (`r sig_label(p_bal_l)`), consistent with the baseline result not being driven by the unbalanced structure.

R3: Alternative lag structures. To assess whether the one-period lag is sufficient to address timing concerns, the full specification is re-estimated using two-year ($t-2$) and three-year ($t-3$) lags of FDI. Each additional lag removes the first year of each country's time series, reducing the sample to `r format(n_obs_l2, big.mark=",")` and `r format(n_obs_l3, big.mark=",")` observations for $t-2$ and $t-3$, respectively. The total below-median-income FDI effects are $`r round(b_l2_l, 3)`$ (`r sig_label(p_l2_l)`) and $`r round(b_l3_l, 3)`$ (`r sig_label(p_l3_l)`), respectively. The gradual attenuation from $t-1$ to $t-3$ is consistent with FDI spillovers materialising within the first couple of years following investment.

R4: Two-way clustered standard errors. Country-only clustering may under-estimate standard errors if unobserved shocks produce cross-sectional correlation within years (e.g., global capital flow surges). Re-clustering on both country and year simultaneously raises the standard error on the above-median-income FDI coefficient to $`r round(se4_h_2w, 3)`$ and on the interaction to $`r round(se4_i_2w, 3)`$, but the total below-median-income effect remains significant at the $`r if(p4_l_2w < 0.001) "0.1\\%" else if(p4_l_2w < 0.01) "1\\%" else if(p4_l_2w < 0.05) "5\\%" else "10\\%"` level.

Despite these checks, three limitations remain. First, the residual endogeneity concern (investors tracking multi-year growth trajectories) cannot be fully resolved without a valid instrument. Bilateral investment treaty ratification networks represent a candidate for future work. Second, the aggregate FDI measure conflates resource-seeking, market-seeking, and efficiency-seeking investment [@dunning1979], as well as reinvested earnings of existing affiliates alongside new greenfield entry; each component carries different implications for technology transfer, and disaggregated bilateral data would allow a more detailed investigation of mechanisms. Third, the income-group indicator $D_i^L$ is constructed from time-averaged GDP per capita over the full sample period, meaning that the growth performance being explained contributes to the group assignment. Countries that grew relatively slowly over the panel period are more likely to be classified as below-median, which could introduce a form of look-ahead bias. The time-varying classification in R1 partially addresses this by using only contemporaneous income, but a fully prospective classification scheme would require external income thresholds.

# Conclusion

This paper examines the relationship between foreign direct investment and GDP per capita growth using an unbalanced panel of `r n_countries` economies over `r yr_min` to `r yr_max`. By combining one-period lagging, two-way fixed effects, and a conditional convergence control, the analysis addresses identification challenges arising from reverse causality and time-invariant omitted variables. Residual endogeneity cannot be ruled out without a valid instrument, so results are best interpreted as conditional associations rather than causal estimates. The central finding is that the aggregate FDI-growth coefficient, while positive, masks income-group heterogeneity. In above-median-income economies, within-country changes in FDI have no statistically significant effect on growth. In below-median-income economies, a one-percentage-point increase in the lagged FDI-to-GDP ratio is associated with approximately `r round(marg, 2)` pp of additional annual GDP per capita growth, an effect that is statistically significant.

These results are consistent with the @findlay1978 relative-backwardness hypothesis: technology spillovers from FDI appear largest where the gap with the global frontier is greatest. The finding is robust to four alternative specifications: replacing the static income-group classification with a time-varying annual-median criterion, restricting estimation to the balanced sub-panel of countries with at least `r min_obs_bal` annual observations, shifting the lag structure to $t-2$ or $t-3$, and tightening inference with two-way country-and-year clustered standard errors. In all cases the total below-median-income FDI effect remains positive and statistically significant.

These results have implications for FDI promotion policy. For above-median-income countries, the estimated growth returns to FDI attraction are negligible on average, suggesting that fiscal incentives aimed at attracting multinationals may offer limited value as a growth strategy. For below-median-income countries, the evidence is consistent with FDI having a positive association with growth, which motivates continued attention to investment-climate conditions. These results cannot, however, identify which specific policies most effectively attract productive FDI or whether the estimated association would persist under policy-induced increases in FDI flows.

Future work should pursue a credible instrumental-variable strategy to sharpen the causal interpretation, decompose the aggregate FDI effect by investment type following the @dunning1979 OLI framework, and examine whether the intensive margin (increased investment by incumbents) and the extensive margin (new entrant multinationals) carry different growth implications.

# References
